---
title: "p8105_hw2_TZ2472"
author: "tianwei zhao"
date: "10/5/2021"
output: github_document
---

```{r,echo=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(haven)
```


## problem 1 
Read and clean the Mr. Trash Wheel sheet:

specify the sheet in the Excel file and to omit non-data entries (rows with notes / figures; columns containing notes) using arguments in read_excel

use reasonable variable names

omit rows that do not include dumpster-specific data

round the number of sports balls to the nearest integer

```{r}
trawheel_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",range = "A2:N408") %>% 
  janitor::clean_names() %>% 
  mutate(sports_balls = round(sports_balls)) %>% 
  drop_na(dumpster)
  
  

  
```



Read and clean precipitation data for 2018 and 2019. For each, omit rows without precipitation data and add a variable for year. 

```{r}
precip2018_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
             sheet =5 ,range ="A2:B15" ) %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(year = "2018")

precip2019_df =
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
             sheet = 4, range = "A2:B15") %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(year = "2019")
  
  
  
```



Next, combine precipitation datasets and convert month to a character variable (the variable month.name is built into R and should be useful).
```{r}
precip1819_df = 
  bind_rows(precip2018_df,precip2019_df) %>% 
  janitor::clean_names() %>% 
  relocate(year) %>% 
  mutate(month = month.name[month]) %>% 
  rename(precipitation = total) #change the variable name total to precipitation for better understanding 
  
  
```

