---
title: "p8105_hw2_TZ2472"
author: "tianwei zhao"
date: "10/5/2021"
output: github_document
---

```{r,echo=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(haven)
```


## problem 1 
Read and clean the Mr. Trash Wheel sheet:

specify the sheet in the Excel file and to omit non-data entries (rows with notes / figures; columns containing notes) using arguments in read_excel

use reasonable variable names

omit rows that do not include dumpster-specific data

round the number of sports balls to the nearest integer

```{r}
trawheel_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",range = "A2:N535") %>% 
  janitor::clean_names() %>% 
  relocate(sports_balls) %>% 
  mutate(sports_balls = round(sports_balls)) %>% 
  drop_na(dumpster)
  
  

  
```



Read and clean precipitation data for 2018 and 2019. For each, omit rows without precipitation data and add a variable for year. 

```{r}
precip2018_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
             sheet = 7 ,range ="A2:B15" ) %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(year = "2018")

precip2019_df =
    read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
             sheet = 6, range = "A2:B15") %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(year = "2019")
  
  
  
```



Next, combine precipitation datasets and convert month to a character variable (the variable month.name is built into R and should be useful).
```{r}
precip1819_df = 
  bind_rows(precip2018_df,precip2019_df) %>% 
  janitor::clean_names() %>% 
  relocate(year) %>% 
  mutate(month = month.name[month]) %>% 
  rename(precipitation = total) #change the variable name total to precipitation for better understanding 
  
  
```


Write a paragraph about these data; you are encouraged to use inline R. Be sure to note the number of observations in both resulting datasets, and give examples of key variables. For available data, what was the total precipitation in 2018? What was the median number of sports balls in a dumpster in 2019?


In the Mr.Trash wheel data set, we have a total of `r nrow(trawheel_df)` observations, each represents a dumpster.The data has `r ncol(trawheel_df)` variables, collects key information about Mr.Trash Weel such as specific dumpster number, date, amount of trash they collected in weight and volumes, types of trash such as bottles, polystyrene, cigarette and bags, and the homes they powered. The median number of sports balls in a dumpster in 2019 is `r trawheel_df %>% filter(year == 2019) %>% pull(sports_balls) %>% median()`.

In the precipitation in 2018 data set, we we have a total of `r nrow(precip2018_df)` observations, and `r ncol(precip2018_df)` variables,which lists out month and precipitation in that month.


In the precipitation in 2019 data set, we we have a total of `r nrow(precip2019_df)` observations, and `r ncol(precip2019_df)` variables,which lists out month and precipitation in that month.

In the combined data set, we have `r nrow(precip1819_df)`observations, and `r ncol(precip1819_df)` variables listing out the month with precipitations in both 2018 and 2019,and also the total amount of precipitation in each month and year. The total precipitation in 2018 is `r precip2018_df %>% filter(year ==2018) %>% pull(total) %>% sum()`.



## Problem 2

First, clean the data in pols-month.csv. Use separate() to break up the variable mon into integer variables year, month, and day; replace month number with month name; create a president variable taking values gop and dem, and remove prez_dem and prez_gop; and remove the day variable.

```{r}
pols_df = read_csv("data/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate("mon",c("year","month","day") ) %>% #break up the variable mon into integer variables year, month, and day
  mutate(month =month.name[as.numeric(month)]) %>% # replace month number with month name
  mutate(president = case_when(
    prez_gop >= 1 ~ "gop",
    prez_dem == 1 ~ "dem")) %>% #create a president variable taking values gop and dem
  select(-c(day,prez_dem,prez_gop))#remove prez_dem and prez_gop; and remove the day variable
  

         
```



Second, clean the data in snp.csv using a similar process to the above. For consistency across datasets, arrange according to year and month, and organize so that year and month are the leading columns.

```{r}
snp_df = read_csv("data/fivethirtyeight_datasets/snp.csv") %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  separate("date",c("year","month","day") ) %>%  
  mutate(year = as.numeric(year),
         month = as.numeric(month),
         year = ifelse(year >2015, year-100,year)) %>% 
  arrange(year,month) %>% 
  mutate (month = month.name[month],
          year = as.character(year))%>% 
  relocate(year,month) # clean the data in snp.csv using a similar process to the above
```


Third, tidy the unemployment data so that it can be merged with the previous datasets. This process will involve switching from “wide” to “long” format; ensuring that key variables have the same name; and ensuring that key variables take the same values.

```{r}
unemploy_df = read_csv("data/fivethirtyeight_datasets/unemployment.csv",
                       col_names = c("year","1":"12"),skip = 1) %>% #rewrite the month to numbers
  pivot_longer( "1":"12",names_to = "month") %>% #arrange months into one colunm which named "month"
  mutate(month = month.name[as.numeric(month)],
         year = as.character(year)) #replace month number with month name
```


Join the datasets by merging snp into pols, and merging unemployment into the result.

```{r}
combined_df = 
  left_join(pols_df, snp_df,by = c("year","month")) %>% 
  left_join(unemploy_df, by = c("year","month"))
```



Write a short paragraph about these datasets. Explain briefly what each dataset contained, and describe the resulting dataset (e.g. give the dimension, range of years, and names of key variables).

pols_df is a 822*9 data frame, which has a range of 68 years, and has variables of year, month,president of either republic or democratic sides,the number of  governors, senators, representatives of either republic or democratic sides on the associated date.

snp_df is a 787*4 data frame, that has a range of 65 years, and has variables of year, month, date, and the closing values of the S&P stock index on the associated date.

unemploy_df is a 816*3 data frame, that has a range of 67 years, and has varibale of year, month, percentage of unemployment in each month of the associated year.

combined_df is a 822*12 data frame, that has a range of 68 years, and has varibale of year, month,president of either republic or democratic sides,the number of  governors, senators, representatives of either republic or democratic sides on the associated date,the closing values of the S&P stock index on the associated date,and percentage of unemployment in each month of the associated year.


## Problem 3

Load and tidy the data. Note that, although these data may seem fairly well formatted initially, the names of a categorical predictor and the case structure of string variables changed over time; you’ll need to address this in your data cleaning. Also, some rows seem duplicated, and these will need to be removed (hint: google something like “dplyr remove duplicate rows” to get started).

```{r}
popular_df = read_csv("data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  mutate(childs_first_name = str_to_lower(childs_first_name),
         gender = str_to_lower(gender),
         ethnicity = str_to_lower(ethnicity)) %>% 
  # make each level ethnicity identical
  mutate(ethnicity = replace(ethnicity, ethnicity == 'asian and paci', 'asian and pacific islander'),
         ethnicity = replace(ethnicity, ethnicity == 'black non hisp', 'black non hispanic'),
         ethnicity = replace(ethnicity, ethnicity == 'white non hisp', 'white non hispanic')) %>% 
  distinct() #remove duplicate rows.

popular_df     
```

Produce a well-structured, reader-friendly table showing the rank in popularity of the name “Olivia” as a female baby name over time; this should have rows for ethnicities and columns for year. Produce a similar table showing the most popular name among male children over time.

```{r}
olivia_female_df = popular_df %>% 
  filter(childs_first_name == 'olivia', gender=='female') %>% 
  select(!c(childs_first_name,gender, count)) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank")
knitr::kable(olivia_female_df, caption = "Rank of the Popularity of Olivia")


most_popular_male_df = popular_df %>% 
  filter(gender == 'male',rank == 1) %>% 
  select(!c(childs_first_name,count)) %>% 
  pivot_wider(names_from = "year_of_birth",
              values_from = "rank")
knitr::kable(most_popular_male_df, caption = "Rank of the Popularity of Male")
  
  
```

Finally, for male, white non-hispanic children born in 2016, produce a scatter plot showing the number of children with a name (y axis) against the rank in popularity of that name (x axis).

```{r}
scatterplot_df = popular_df %>% 
  filter(gender == 'male', ethnicity == 'white non hispanic', year_of_birth == 2016)

ggplot(scatterplot_df, aes(x=rank, y=count)) +
  geom_point(color = 'lightblue') + 
  ggtitle("Scatter plot of white non hispanic male children born in 2016") +
  ylab("Number of children with that name") + 
  xlab("Rank in popularity of one name")
```



